Bulk Semantic Annotation with an Partially Known Knowledge Base

### 2024-11-12 doc/cicm25-sysdesc/sysdesc.tex
more

Concretely this is about converting the {\LaTeX} string
\begin{lstlisting}[language={[LaTeX]TeX}]
\item \blue{Problem:} The state space has local minima, e.g. the board on the
  right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
into the (partially annotated item:  
\begin{lstlisting}[language={[LaTeX]TeX},morekeywords={sn,importmodule}]
\importmodule[smglom/search]{mod?state-space}
[...]  
\item \blue{Problem:} The \sn{state space} has local minima, e.g. the
  board on the right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
Concretely, the author has to remember the module for the term ``state space'' from
\cref{fig:state-space}, the symbol name \lstinline|state space|, annotate it with the
\sTeX macro \lstinline|\sn{...}|, and add the module import in the first line (unless it
is already imported). In this case the symbol name happens to be the same as its English
verbalization, otherwise we would have to use
\lstinline[mathescape]|\sr{$\llangle symbol name\rrangle$}{state space}|.

@@

The \srf system is a simple command-line tool that creates a database of
symbol-verbalization pairs\ednote{explain that first}, analyzes \sTeX source files and
then steps through all word occurrences in the document that share the stem with a
verbalization in the database. For each such word, the user is presented with an
annotation choice and interactions that allow to fine-tune the wordwise annotation
workflow.

@@

In this system description we have presented the \srf tool, a command-line tool

>>>>

Concretely this is about converting the {\LaTeX} string\ednote{add numbers, start with 3
  so that corresponding lines with the next listing.}
\begin{lstlisting}[language={[LaTeX]TeX}]
\item \blue{Problem:} The state space has local minima, e.g. the board on the
  right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
into the (partially annotated item:
\begin{lstlisting}[language={[LaTeX]TeX},morekeywords={sn,importmodule}]
\importmodule[smglom/search]{mod?state-space}
[...]  
\item \blue{Problem:} The \sn{state space} has local minima, e.g. the
  board on the right has $h=1$ but every successor has $h>1$. 
\end{lstlisting}
Concretely, the author has to remember the module for the term ``state space'' from
\cref{fig:state-space}, the symbol name \lstinline|state space|, annotate it with the
\sTeX macro \lstinline|\sn{...}|, and add the module import in the first line (unless it
is already imported). In this case the symbol name happens to be the same as its English
verbalization, otherwise we would have to use
\lstinline[mathescape]|\sr{$\llangle symbol name\rrangle$}{state space}|.

@@

The \srf system (see \cite{stextools:git} is a simple command-line tool that creates a
database of symbol-verbalization pairs\ednote{explain that first}, analyzes \sTeX source
files and then steps through all word occurrences in the document that share the stem with
a verbalization in the database. For each such word, the user is presented with an
annotation choice and interactions that allow to fine-tune the wordwise annotation
workflow.

@@

In this system description we have presented the \srf tool, a command-line tool
\ednote{continue}

In an earlier attempt to support semantic annotation in IDEs, we tried named-entity
recognition (NER) for classification of ``likely annotation candidate words''
\cite{hutterer:msc23}, however this classification was not precise enough in the
distinction in ``technical terms'' and ordinary english noun phrases and named entities --
the relevant task for annotation, and so made it impractical, especially, since it could
not enter the actual annotations and import directives automatically. But maybe using the
NER-based approach, after the \srf annotation process might change the tradeoffs
involved. In any case, the overall workflow suggested by the NER-based approach --
especially after the verbalizations covered by the existing content commons\ednote{better
  to speak of the domain model here?} -- is geared towards adding to the content commons.

This is complementary to the \srf approach,
